Sparse Function-space Representation \\ of Neural Networks

### 2023-05-13 paper/main.tex
written some dual variable stuff

\textbf{Sparse Dual Gaussian Process} Sparse Gaussian processes are a way to reduce the computational complexity of a full GP using a low-rank approximation. Many sparsifying approaches exist (see \citet{quinonero2005unifying}) for an overview. Given that we have our dual variables derived from our neural network predictions and kernel function, we could use any approximation. We decide to follow the sparsifying approach of \citep{titsias2009variational}, which is to define the marginal distribution as $q_{\vu}(\vf)  = \int p(\vf  \mid \vu) q(\vu)  d\vu$. Given that we have $p(f_i  \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$. As was shown in \cite{adam2021dual}, the posterior under this model has a particular structure similar to \cref{eq:gp_pred}. In that paper, they spot the dual variables' importance for approximate inference but write them in the natural parameter form as this form is more suitable for optimization through natural gradients. In order to make the link to the dual variables defined in \cref{eq:dual_param} we write them in the sparse GP in the dual variable form.

\begin{equation} 
	\valpha_{\vu}  = \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \hat{\alpha}_{i}, \quad
	\vbeta_{\vu} = \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}

Sparse Gaussian processes \citep{hensman2013gaussian} are a principled way to scale Gaussian processes to large datasets by using a subset of $m$ inducing points, with $m \ll n$,  lowering the complexity from $\mathcal{O}(n^3)$ to $\mathcal{O}(m^3)$. The key insight is to use a sparse approximating distribution $q(\vu; \vm, \MV)$, where $\vm \in \R^{m}$ and $\MV \in \R^{m \times m}$, and find its parameters through maximizing the evidence lower bound (ELBO). Then, the distribution over all points is defined as $q_{\vu}(\vf) = \int p(\vf \mid \vu) q(\vu) \mathrm{d}\vu$. As shown in \cite{adam2021dual}, we can decompose the optimal sparse GP  with a dual parameterization \cite{adam2021dual} as follows:
\begin{equation}
   \myexpect_{q_{\vu}(\vf)}[\vf_i] = \vkzs^{\T} \MKzz^{-1}  \valpha_{\vu} , \quad 
   \textrm{Var}_{q_{\vu}(\vf)}[\vf_i]  = \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \vbeta_{\vu})^{-1} ]\vkzs
\end{equation}

where the $ij$th entry of the matrix $\MKzz \in \R^{m \times m}$ is $\kappa(\vz_i,\vz_j)$, $\vk_{\vz i}$\todo{the equation above uses $\vkzs$ not $\vk_{\vx i}$} denotes a vector where each $j$th element is $\kappa(\vz_j, \vx_i)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and the matrix $\vbeta \in \R^{m \times m}$, defined as
\begin{equation} 
	\vlambda = [ \MKzz + \vbeta] \valpha  = \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \tilde{\lambda}_{i}, \quad
	\vbeta = \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}

>>>>

\textbf{Sparse Dual Gaussian Process} Sparse Gaussian processes are a way to reduce the computational complexity of a full GP using a low-rank approximation. Many sparsifying approaches exist (see \citet{quinonero2005unifying}) for an overview. Given that we have our dual variables derived from our neural network predictions and kernel function, we could use any approximation. We decide to follow the sparsifying approach of \citep{titsias2009variational}, which is to define the marginal distribution as $q_{\vu}(\vf)  = \int p(\vf  \mid \vu) q(\vu)  d\vu$. Given that we have $p(f_i  \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$. As was shown in \cite{adam2021dual}, the posterior under this model has a particular structure similar to \cref{eq:gp_pred}. In that paper, they spot the dual variables' importance for approximate inference but write them in the natural parameter form as this form is more suitable for optimization through natural gradients. In order to make the link to the dual variables defined in \cref{eq:dual_param} we write them in the sparse GP in the dual variable form. 
\begin{equation} 
	\valpha_{\vu}  = \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \hat{\alpha}_{i}, \quad
	\vbeta_{\vu} = \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}
Using the sparse definition of the dual variables our sparse GP posterior takes the following form:
\begin{equation}
   \myexpect_{q_{\vu}(\vf)}[\vf_i] = \vkzs^{\T} \MKzz^{-1}  \valpha_{\vu} , \quad 
   \textrm{Var}_{q_{\vu}(\vf)}[\vf_i]  = \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \vbeta_{\vu})^{-1} ]\vkzs
\end{equation}

where $\MKzz$ and $\vkzs$ are define similarly to $\MKxx$ and $\vk_{\vx i}$ but now over the $\MZ \in \R^{M \times D} $ inducing points instead of the full dataset.

Sparse Gaussian processes \citep{hensman2013gaussian} are a principled way to scale Gaussian processes to large datasets by using a subset of $m$ inducing points, with $m \ll n$,  lowering the complexity from $\mathcal{O}(n^3)$ to $\mathcal{O}(m^3)$. The key insight is to use a sparse approximating distribution $q(\vu; \vm, \MV)$, where $\vm \in \R^{m}$ and $\MV \in \R^{m \times m}$, and find its parameters through maximizing the evidence lower bound (ELBO). Then, the distribution over all points is defined as $q_{\vu}(\vf) = \int p(\vf \mid \vu) q(\vu) \mathrm{d}\vu$. As shown in \cite{adam2021dual}, we can decompose the optimal sparse GP  with a dual parameterization \cite{adam2021dual} as follows:

denotes a vector where each $j$th element is $\kappa(\vz_j, \vx_i)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and the matrix $\vbeta \in \R^{m \times m}$, defined as
\begin{equation} 
	\vlambda = [ \MKzz + \vbeta] \valpha  = \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \tilde{\lambda}_{i}, \quad
	\vbeta = \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}
